A Simple Ratio for Estimating the Effective Reproduction Value of an Epidemic

### 2021-01-16 docs/manuscript/manuscript.Rmd
changed referencing to cref package

---
title: "A Simple Ratio for Estimating the Effective Reproduction Value of an Epidemic"
authors:
- name: Authors
  email: emails
  address: institutions
abstract: |
  ABSTRACT
acknowledgements: |
  Acknowledgements
output:
  bookdown::pdf_book:
    base_format: rticles::oup_article
    number_sections: no
    extra_dependencies: float
    keep_tex: yes
    includes:
      in_header: header.tex
  bookdown::word_document2:
    number_sections: no
    extra_dependencies: float
    keep_tex: yes
    includes:
      in_header: header.tex
bibliography:
- packages.bib
- references.bib
keywords:
- key
- dictionary
- word
---

@@

To assess accuracy of $R_t$ estimations, we used Susceptible-Exposed-Infectious-Resolved (SEIR) model to simulate epidemics and then compare how accurately the two methods can estimate the $R_t$ values in the epidemics. We simulated `r simulation_parameters$Nsims` epidemics in a population of `r addUnits(N)` individuals. We varied the lengths of latent and infectiousness periods with each simulation. In each simulation, the population underwent changes in the $R_t$ value representing changes in population behavior (e.g. "lockdown" period or lifting of a lockdown). **Figure** \@ref(fig:sim-results) **A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic. A full visualization of the all the simulations can be seen in (**Figure** \@ref(fig:all-simulations)).

Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (**Figure** \@ref(fig:sim-results) **B**). The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (**Figure** \@ref(fig:sim-results) **C**). This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].

@@

Often however, the the mean and standard deviation of the generation interval is typically not known exactly. For instance, for the COVID-19 the mean generation interval estimates, with 95% credible intervals, include 3.95 (3.01 - 4.91) and 5.20 (3.78 - 6.78); and estimates of the standard deviation of the generation interval include 1.51 (0.74 - 2.97) and 1.72 (0.91 - 3.93) [@ganyani_estimating_2020]. Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer. Under these conditions, the simple ratio actually performs slightly better than than the Cori method (**Figure** \@ref(fig:estimation-inaccParams)).

@@

We next compared $R_t$ estimates from the Simple Ratio and Cori et al for using data from the current COVID-19 pandemic. We estimated $R_t$ in four nations using a generation interval with mean of `r COVID19_params$mean_gi` and standard deviation of `r COVID19_params$sd_gi`[@ganyani_estimating_2020]. **Figure** \@ref(fig:real-world-data) **A** shows the results for the month of October 2020. Visual inspection shows that the two methods of estimating $R_t$ have very good concordance. Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (**Figure** \@ref(fig:real-world-data) **B**).

>>>>

---
title: "A Simple Ratio for Estimating the Effective Reproduction Value of an Epidemic"
authors:
- name: Authors
  email: emails
  address: institutions
abstract: |
  ABSTRACT
acknowledgements: |
  Acknowledgements
output:
  bookdown::pdf_book:
    base_format: rticles::oup_article
    number_sections: no
    extra_dependencies: float
    keep_tex: yes
    includes:
      in_header: header.tex
  bookdown::word_document2:
    number_sections: no
    extra_dependencies: float
    keep_tex: yes
    includes:
      in_header: header.tex
header-includes:
  - \usepackage[labelsep=period,labelfont=bf]{caption}
  - \usepackage{hyperref}
  - \usepackage[capitalize, nameinlink]{cleveref}
  - \crefdefaultlabelformat{#2\textbf{#1}#3} % <-- Only #1 in \textbf
  - \crefname{figure}{\textbf{Figure}}{\textbf{Figures}}
  - \Crefname{figure}{\textbf{Figure}}{\textbf{Figures}}
  - \crefname{table}{\textbf{Table}}{\textbf{Tables}}
  - \Crefname{table}{\textbf{Table}}{\textbf{Tables}}
bibliography:
  - packages.bib
  - references.bib
keywords:
  - key
  - dictionary
  - word
---

@@

To assess accuracy of $R_t$ estimations, we used Susceptible-Exposed-Infectious-Resolved (SEIR) model to simulate epidemics and then compare how accurately the two methods can estimate the $R_t$ values in the epidemics. We simulated `r simulation_parameters$Nsims` epidemics in a population of `r addUnits(N)` individuals. We varied the lengths of latent and infectiousness periods with each simulation. In each simulation, the population underwent changes in the $R_t$ value representing changes in population behavior (e.g. "lockdown" period or lifting of a lockdown). \cref{fig:sim-results}**A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic. A full visualization of the all the simulations can be seen in (\cref{fig:all-simulations)).

Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results)**B**). The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results)**C**). This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].

@@

Often however, the the mean and standard deviation of the generation interval is typically not known exactly. For instance, for the COVID-19 the mean generation interval estimates, with 95% credible intervals, include 3.95 (3.01 - 4.91) and 5.20 (3.78 - 6.78); and estimates of the standard deviation of the generation interval include 1.51 (0.74 - 2.97) and 1.72 (0.91 - 3.93) [@ganyani_estimating_2020]. Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer. Under these conditions, the simple ratio actually performs slightly better than than the Cori method (\cref{fig:estimation-inaccParams)).

@@

We next compared $R_t$ estimates from the Simple Ratio and Cori et al for using data from the current COVID-19 pandemic. We estimated $R_t$ in four nations using a generation interval with mean of `r COVID19_params$mean_gi` and standard deviation of `r COVID19_params$sd_gi`[@ganyani_estimating_2020]. \cref{fig:real-world-data)**A** shows the results for the month of October 2020. Visual inspection shows that the two methods of estimating $R_t$ have very good concordance. Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (\cref{fig:real-world-data)**B**).
